---
title: "Oxygen Saturation off California Coast Models"
author: "Margaret Hammond"
date: "1/21/2022"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(AICcmodavg)
library(tidyverse)
library(janitor)
library(here)
library(equatiomatic)
library(kableExtra)

```


### Overview
In this task, I will explore the relationship betwen oxygen saturation of seawater off of California's coast with several physical and chemical variables. This data comes from a 70+ year hydrographic time-series. I will be comparing two models with AIC and mean RMSE. 


#### Read in the Data

```{r}
# Read the data

calcofi <- read_csv(here("data", "calcofi_seawater_samples.csv"))
```


#### Compare the Models
Compare the two linear models of oxygen saturation by their AIC values and analyzing the results. Model 1 focuses on oxygen as a function of water temperature (C), salinity (Practical Salinity Scale), and phosphate concentration (micro Moles per Liter).

```{r}
# Comparison of two models using an appropriate application of AIC, with the resulting AIC scores and an interpretation of how to choose the best model with this information

calcofi_clean <- calcofi %>% 
  rename(temp = t_deg_c,
         phosph = po4u_m,
         depth = depth_m)
# Oxygen saturation (o2sat) as a function of water temperature (t_deg_c), salinity (salinity), and phosphate concentration (po4u_m)
f1 <- o2sat ~ temp + salinity + phosph
mdl1 <- lm(o2sat ~ temp + salinity + phosph, data = calcofi_clean)


# Oxygen saturation (o2sat) as a function of water temp, salinity, phosphate concentration, and depth 

f2 <- o2sat ~ temp + salinity + phosph + depth
mdl2 <- lm(o2sat ~ temp + salinity + phosph + depth, data = calcofi_clean)

# Compare the AIC

#AIC(mdl1, mdl2)
#aictab(list(mdl1, mdl2))

aictab(list(mdl1, mdl2)) %>% 
  kable(col.names = c("Model Name",
                      "DF",
                      'AICc',
                      'Delta AIC',
                      'ModelLik',
                      'AICcWt',
                      'LL',
                      'Cum.Wt'),
        caption = 'Table 1. Comparison of AIC values for Model 1 and 2',
        digits = 2) %>% 
  kable_styling(full_width = FALSE)
```


#### AIC Choice
From the AIC test run, model 2 containing a model of oxygen saturation as a function of water temp, salinity, phosphate concentration, and depth is a better choice than model 1. Model 2 has an AIC value of 616.60, compared to Model 1's AIC of 619.03. Model 2 is 2 points lower than Model 1, making it a significantly better model. 


### K-fold cross validation
Compare the two linear models using a K-fold cross validation, with resulting average RMSE value across all folds. 
```{r}


# Creating the 10-fold cross validation
folds <- 10
fold_vec <- rep(1:folds, length.out = nrow(calcofi_clean))
table(fold_vec)

set.seed(41)

calcofi_fold <- calcofi_clean %>% 
  mutate(group = sample(fold_vec, 
                        size = n(),
                        replace = FALSE))

test_df <- calcofi_fold %>% 
  filter(group == 1)

train_df <- calcofi_fold %>% 
  filter(group != 1)

# RMSE calculation

calc_rmse <- function(x, y) {
  rmse_result <- (x - y)^2 %>% 
    mean() %>% 
    sqrt()
  return(rmse_result)
}

```


#### Use the training dataset
USe the training dataset to create 2 linear regression models, based on the formulas above.
```{r}
# Use the training dataset to create 2 linear regression models, based on the formulas above

training_mdl1 <- lm(f1, data = train_df)
training_mdl2 <- lm(f2, data = train_df)

# Use trained models to predict on test data

predict_test <- test_df %>% 
  mutate(model1 = predict(training_mdl1,
                          test_df),
         model2 = predict(training_mdl2, test_df))

rmse_predict_test <- predict_test %>% 
  summarize(rmse_mdl1 = calc_rmse(model1, o2sat),
            rmse_mdl2 = 
              calc_rmse(model2, o2sat))

rmse_predict_test %>% 
 kable(col.name = c('RMSE Model 1',
                    'RMSE Model 2'),
       caption = 'Table 2. RMSE of 1 Interation of k-fold for Model 1 and 2',
       digits = 2) %>% 
  kable_styling(full_width = FALSE)


```



#### 10-fold cross validation 
Create a 10-fold cross validation across all folds to calculate average RMSE.
```{r}
# Calculate over all the folds

rmse_df <- data.frame()

for(i in 1:folds) {
  kfold_test_df <- calcofi_fold %>% 
    filter(group == i)
  kfold_train_df <- calcofi_fold %>% 
    filter(group != i)
  
  kfold_mdl1 <- lm(f1, data = kfold_train_df)
  kfold_mdl2 <- lm(f2, data = kfold_train_df)
  
  kfold_pred_df <- kfold_test_df %>% 
    mutate(mdl1 = predict(kfold_mdl1, kfold_test_df),
           mdl2 = predict(kfold_mdl2, .))
  kfold_rmse <- kfold_pred_df %>% 
    summarize(rmse_mdl1 =
                calc_rmse(mdl1, o2sat),
              rmse_mdl2 =
                calc_rmse(mdl2, o2sat))
  rmse_df <- bind_rows(rmse_df, 
                       kfold_rmse)
}

rmse_sum <- rmse_df %>% 
  summarize(mean_rmse_mdl1 =
              mean(rmse_mdl1),
            mean_rmse_mdl2 =
              mean(rmse_mdl2))

rmse_sum %>% 
   kable(col.name = c('RMSE Model 1',
                    'RMSE Model 2'),
       caption = 'Table 3. RMSE of 10 Interations of k-fold for Model 1 and 2',
       digits = 2) %>% 
  kable_styling(full_width = FALSE)

```


### Model Choice by Average RMSE
Model 2 is chosen as a better overall linear model due to it's smaller mean RMSE value. Model two is oxygen saturation as a function of water temp, salinity, phosphate concentration, and depth. 

```{r}
# Choosing model 2 because it has the smaller mean RMSE. 


final_mdl <- lm(f2, data = calcofi_clean)
sum_mdl1 <- summary(final_mdl)
```


### Final Model

`r extract_eq(final_mdl, wrap = TRUE, use_coefs = TRUE)`


### Citation
CalCOFI data are available for use without restriction. Data downloaded from https://calcofi.org/ccdata.html.  Accessed 1/10/2022.
