---
title: "Mountain Yellow-Legged Frog in Sierra, Nevada"
author: "Margaret Hammond"
date: "1/18/2022"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(readxl)
library(lubridate)
library(patchwork)

```


### Overview
This task is focused on exploring data containing information on amphibian abundance in the Sierra Lakes Inventory Project (SLIP). The SLIP ecosystem contains aquatic communities of >8,000 lentic water bodies in the southern Sierra Nevada. In this task, we are focusing on the mountain yellow-legged frog (*Rana Muscosa*). In part one, a graph is created to visualize the different population size for life stages of mountain yellow-legged frog (*Rana Muscosa*). In opart two, a column graph is created containing total counts of combined adult and subadult mountain yellow-legged frog (*Rana Muscosa*). 


#### Read in the Data
```{r}
amphib <- read_xlsx(here("data", "sierra_amphibians.xlsx"))
```



#### Wrangle the Data
Wrangle the data to create a plot of mountain yellow-legged frog count each year across all waster bodies by life stage, excluding the `EggMass` stage.

```{r}


amphib_date <- amphib %>% 
  drop_na() %>%
  filter(amphibian_species == "RAMU") %>% 
  filter(amphibian_life_stage != "EggMass") %>% 
  mutate(survey_date = ymd(survey_date)) %>% 
  mutate(year = year(survey_date)) %>% 
  group_by(year, amphibian_life_stage) %>% 
  summarize(amphib_sum = sum(amphibian_number))

p1 <- ggplot(data = amphib_date, aes(x = year, y = amphib_sum)) +
  geom_bar(aes(fill = amphibian_life_stage), stat = "identity") +
  theme_minimal() +
  labs(x = "Year",
       y = "Amount of Frogs",
       title = "Mountain Yellow-Legged Frog Population Over Time",
       fill = 'Life Stage',
       caption = "Figure 1. Mountain yellow-legged frog population 
       in the Sierra, Nevada lakes grouped by lifestage from 1995 - 2002.")

p1

```

```{r}
# A column graph containing total counts (over all years of the study) of combined adult and subadult endangered mountain yellow-legged frogs (Rana muscosa) observed in the 5 lakes with the greatest total observed counts


amphib_lake <- amphib %>% 
  filter(amphibian_species == "RAMU") %>% 
  filter(amphibian_life_stage %in% c("Adult", "SubAdult")) %>% 
  group_by(lake_id) %>% 
  summarize(amphib_n = sum(amphibian_number)) %>% 
  slice_max(amphib_n, n = 5) %>% 
  mutate(lake_id = paste('Lake', lake_id, sep = ' ')
  ) %>% 
  mutate(lake = fct_reorder(lake_id, amphib_n))
  
p2 <- ggplot(data = amphib_lake, aes(x = lake, y = amphib_n)) +
  geom_bar(stat = 'identity', fill = 'green4') +
  theme_minimal() +
  labs(x = 'Lake ID',
       y = 'Number of Frogs',
       title = 'Endangered Mountain Yellow Frogs Population by Lake',
       caption = 'Figure 2. Top 5 Lakes in Sierra, Nevada with highest 
       adult and subadult endangered mountain 
       yellow-legged frogs populations from 1995-2002.')

p2
```


#### Combine with Patchwork
```{r}
# Combine the two plots with patchwork

p1 / p2


```

### Citation
Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3
